### 2018-11-21 response/response-demography-1.tex

\documentclass[letterpaper,12pt]{article}

% Xetex preamble
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}

\usepackage[longnamesfirst]{natbib}
%\usepackage[tabhead,nolists,tablesfirst]{endfloat}
\usepackage[flushleft]{threeparttable}
\usepackage{booktabs}
\usepackage{rotating}  
\usepackage{dcolumn} 
\usepackage{setspace}
%\usepackage{flafter} 
%\usepackage{longtable}
%\usepackage[pdftex]{graphicx}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black]{
hyperref} 
\usepackage[margin=1.0in]{geometry} 
\usepackage{multirow}

%opening
\title{} \author{}

\doublespacing

\begin{document}

\begin{center} \textbf{\large Birth Spacing in the Presence of Son
Preference and 
Sex-Selective Abortions: India's Experience over Four Decades}
\end{center}

\begin{center} Response to Editor and Referee Comments \end{center}

\noindent Please find attached the revised version of my paper,
``Birth Spacing in the Presence of Son Preference and Sex-Selective
Abortions:
India's Experience over Four Decades.''
I would like to thank the two referees and the editor for the very
useful comments and
suggestions, which I believe have substantially improved the paper.

[Comments from editor]

First, the description of the problem and literature review are sparse.
The front end of the paper should be revised to situate the study in the
wider literature on son preference and sex selection in the region. The
contributions of the current study should be highlighted. R1 offered
particularly useful ideas in this regard. These changes would make the
paper appeal to the broad readership of Demography.

Second, a conceptual framework or theoretical argument is needed to
motivate the decision to examine birth intervals and sex composition by
maternal education. Does maternal education proxy for access to sex
selective technology, preferences for sons, etc.? In addition, how does
selection into the highest/lowest educated categories change over the 45
years of the study? How could this impact the interpretation of results?

Third, the paper is heavy on the description of methods and results.
These sections could be shortened and reorganized to focus on the most
important issues. R1 also offered suggestions regarding the presentation
of findings.

Fourth, discussion of the fertility transition is neglected and only
discussed in the conclusion. At a minimum, you should consider providing
details on fertility levels and distributions across parities over time.
Is there an easy way to show the sample sizes/proportion of births in
each category of sex composition across years in tables 1-3? More could
be said about the potential impact of the ongoing fertility decline.

\newpage

\section*{Reviewer 1}

1. The key contribution of the paper is the use of birth spacing and sex
ratios to examine changes and differentials in possible sex-selective
abortions. While there is large literature using differential stopping
behaviours (DSB) and sex ratios to examine sex selection, the use of
birth spacing is new and could potentially provide new insights. The
methodology used is rigorous and suited for the purposes of the paper.

2. The paper would need to articulate its contribution clearly. The
paper seems to be focussed too much on methodology and less on
contextual, theoretical and substantive matters. As there is already
well established literature on sex-selection and son preference, the
paper must articulate what is substantially different in the paper. As
one of the key features is the focus on spacing, expanded review of the
literature on spacing is required.

3. The literature review must be systematic. The paper seems to move too
quickly to estimation strategy. Before presenting this, the literature
on spacing, sex ratios, sexselection should be presented. This would
help the paper to make a stronger case of its place and contribution to
the demographic literature.

4. The demographic literature, for example, by works from Christophe Z
Guilmoto, doesn’t seem to be consulted. Guilmoto and colleagues many
papers including recent work in Population Studies and Lancet are
important to consult as they cover the same substantial issues of sex
ratios, sex selection and abortions as addressed in this paper. Though
these demographic papers do not necessarily use spacing, a similar
understanding of sex-selection could be obtained using examining parity
progression. Therefore, the added insights from using spacing need to
articulated much more forcefully.

5. The use of competing hazard models is not new in demographic research
and doesn’t require long exposition. This section can be cut
sustainably. I don’t think the methodological innovation is a key
contribution of the paper as the methods are established even if they
have not been used to examine spacing.

6. The use of NFHS data including the most recent data are welcome. An
additional limitation to those addressed in the paper is the issue of
date imputation in DHS/NFHS. NFHS as with other DHS data routinely use
imputations for year and month of birth. I don’t have the reference for
India (please check data quality reports on DHS website), but the
imputation of month of birth was done for large number of cases in
earlier rounds of DHS. This could be a possible reason for seeming large
number of pre-marital conceptions which do not make sense in the context
of India.

7. It is not clear why exactly 22, 23, 25 years are used a cut-off
points.

8. I understand the empirical reason for not increasing the number of
variables. However, a key variable that is essential to consider is
region. This is important because regional influences on birth spacing
and sex selection are very strong as established in the literature. Sex
selection and abortion is especially strong in certain regions but not
in others. The regional differentials in fertility transition,
contraceptive use, spacing, sex ratios, abortions are clearly documented
in the literature. So it is vital to include this.

9. The three tables are logically organized and contain a wealth of
information. But there are too many comparisons to make in each of the
tables and I wonder if it would make sense to split the tables or
present some of the information in a much more reader friendly way (for
instance by using charts for some of the information).

10. As sex composition and sex ratios has been examined in the literature, I think
it would make sense to not present findings on that in great detail. As
the paper states, the central question addressed is spacing, and so all
other questions should not be dealt in detail

11. The results talk about compression of spacing. This is important and
literature on it should be presented in the intro or lit review section.
There is evidence of compression of the reproductive span at least in
some regions (see Padmadas, et. Al 2004). It would be important to
consider how this fits in with the results presented.

12. As there are many results presented, it would make sense to have a
paragraph at the beginning of the conclusion highlighting the key
findings.

13. Overall, the paper is empirically strong but needs to build on the
contextual, theoretical, and prior demographic literature. As the
readers of Demography are diverse, it would be good to have a balanced
presentation of all the sections. Currently it seems the paper is too
concerned with the methodology and empirical work at the expense of
other areas.

\newpage

\section*{Reviewer 7}

The article was a careful, innovative, and thorough piece of research
picking up on a conundrum concerning how son preference and sex
selection affect birth intervals. I very much enjoyed reading this piece
and think it has useful applications outside of India.  My review is
short as I do not see much to be improved upon.  My only (minor)
suggestion concerns the use of the word "spell".  It is very clear what
you mean by a spell when the paper explains this, but I think that it is
somewhat confusing to put this in the abstract without an explanation. 
It was not until I got to the explanation in the methods that I
understood the abstract fully.  I suggest rewording this section of the
abstract so it is clear what is meant by spell.

\newpage
\bibliographystyle{econometrica}
\bibliography{../paper/sex_selection_spacing.bib}
\addcontentsline{toc}{section}{References}

\end{document}

### 2018-11-23 response/response-demography-1.tex

\begin{document}

\begin{center} \textbf{\large Birth Spacing in the Presence of Son
Preference and 
Sex-Selective Abortions: India's Experience over Four Decades}
\end{center}

\begin{center} Response to Editor and Referee Comments \end{center}

\noindent Please find attached the revised version of my paper,
``Birth Spacing in the Presence of Son Preference and Sex-Selective
Abortions:
India's Experience over Four Decades.''
I would like to thank the two referees and the editor for the very useful comments and 
suggestions, which I believe substantially improved the paper.

\section*{Editor}

@@

First, the description of the problem and literature review are sparse.
The front end of the paper should be revised to situate the study in the
wider literature on son preference and sex selection in the region. The
contributions of the current study should be highlighted. R1 offered
particularly useful ideas in this regard. These changes would make the
paper appeal to the broad readership of Demography.

[Response:]

- Spacing as an indicator for son preference
- The relationship between spacing and child health
- The relationship between spacing and women's health
- Important for understanding how spacing in general behaves. See last point from
editor: if lower fertility, better health, and increased female labor market participation 
are all is associated with a compression of spacing, the use of sex selection would
work against this trend and make it look like India is not behaving like other countries.
- Women's role in general

@@

Third, the paper is heavy on the description of methods and results.
These sections could be shortened and reorganized to focus on the most
important issues. R1 also offered suggestions regarding the presentation
of findings.

[Response:]

Fourth, discussion of the fertility transition is neglected and only
discussed in the conclusion. At a minimum, you should consider providing
details on fertility levels and distributions across parities over time.
Is there an easy way to show the sample sizes/proportion of births in
each category of sex composition across years in tables 1-3? More could
be said about the potential impact of the ongoing fertility decline.

[Response:]

[Generically, spacing tend to decline with lower fertility (check this). 
Presumably the decline in spacing is because part of the cause is higher female labor 
market participation, which makes it more attractive to space children closer together.
Not clear whether this is the case for India - The R1 suggested paper might be a good
starting point with a tie-in to compression.
Compression can also come from improved health and/or less migration.

On the flip side, lower fertility place additional pressure on women to use
sex selection, which lead to *less* compression.
]

@@

[Response:]

[See response to Editor]

4. The demographic literature, for example, by works from Christophe Z
Guilmoto, doesn’t seem to be consulted. Guilmoto and colleagues many
papers including recent work in Population Studies and Lancet are
important to consult as they cover the same substantial issues of sex
ratios, sex selection and abortions as addressed in this paper. Though
these demographic papers do not necessarily use spacing, a similar
understanding of sex-selection could be obtained using examining parity
progression. Therefore, the added insights from using spacing need to
articulated much more forcefully.

[Response:]

[I should have those somewhere already. I think they were in the original version.]

@@

5. The use of competing hazard models is not new in demographic research
and doesn’t require long exposition. This section can be cut
sustainably. I don’t think the methodological innovation is a key
contribution of the paper as the methods are established even if they
have not been used to examine spacing.

[Response:]

[Still need to include the sex ratio calculations it would seem.
Prior readers have asked for those.]

@@

6. The use of NFHS data including the most recent data are welcome. An
additional limitation to those addressed in the paper is the issue of
date imputation in DHS/NFHS. NFHS as with other DHS data routinely use
imputations for year and month of birth. I don’t have the reference for
India (please check data quality reports on DHS website), but the
imputation of month of birth was done for large number of cases in
earlier rounds of DHS. This could be a possible reason for seeming large
number of pre-marital conceptions which do not make sense in the context
of India.

[Response:]

[Maybe run regression only on those birth intervals that does not
involve imputations on either end of the spacing.
This could be presented in the Appendix, unless very different from
results.]

@@

[Response:]

[Add this discussion back into Appendix]

8. I understand the empirical reason for not increasing the number of
variables. However, a key variable that is essential to consider is
region. This is important because regional influences on birth spacing
and sex selection are very strong as established in the literature. Sex
selection and abortion is especially strong in certain regions but not
in others. The regional differentials in fertility transition,
contraceptive use, spacing, sex ratios, abortions are clearly documented
in the literature. So it is vital to include this.

[Response:]

[Check for broad regions to incorporate. Check how to do this for the non-proportional
hazard specification.]

@@

10. As sex composition and sex ratios has been examined in the literature, I think
it would make sense to not present findings on that in great detail. As
the paper states, the central question addressed is spacing, and so all
other questions should not be dealt in detail

[Response:]

11. The results talk about compression of spacing. This is important and
literature on it should be presented in the intro or lit review section.
There is evidence of compression of the reproductive span at least in
some regions (see Padmadas, et. Al 2004). It would be important to
consider how this fits in with the results presented.

[Response:]

[Compression is only for first birth as far as I remember. Need to check that.
Check on paper for their analysis.]

### 2018-11-28 paper/sex_selection_spacing-ver2.tex
Copy of original submission for editing
Update month
First go at reducing length of Estimation Strategy section
Based on Editor's and R1's comments, trying to make the discussion
of the estimation strategy shorter. Still need to highlight how
the approach is different.

\date{March 2018\\
\bigskip
Preliminary}

@@

The unit of analysis is a spell, the period from one birth, or marriage for
the first spell, to the next.
For some spells, the duration is censored, for example, because the survey
took place before the birth that would end the spell or because the couple
has finished childbearing.
The standard approach to address the censoring of spells, which I follow here, is 
to use a hazard model.
To capture the effects of sex selection, I extend the standard hazard model 
to allow for multiple exit states and a non-proportional hazard specification.

First, I use a competing risk framework with two possible exit states, either a boy
or a girl is born, because access to prenatal sex determination means that
parents can choose the sex of children born.
Prior research on birth spacing, in contrast, relied on single
exit state hazard models.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}
Having only one exit state works when there is no sex selection---making the 
sex of the child a random event---but not here since at least some couples use 
sex selection.

Second, I use a non-proportional hazard specification because
sex composition and the use of sex selection are likely to affect 
the shape of the hazard functions across groups.
The use of a non-proportional specification also mitigates any potential 
effects of unobserved heterogeneity when used in conjunction with a 
flexible baseline hazard \citep{Dolton1995}.
Proportional hazard models, where covariates have a multiplicative 
effect on the hazard rate, are more efficient than non-proportional 
models, but only provided that the proportionality assumption hold.
If the proportionality assumption does not hold, the estimates
are biased.

It is unlikely---even in the absence of prenatal sex determination---that 
the effect of, for example, the sex composition of previous births have 
the same effect throughout the entire spell.
Assuming that the effect of sex composition is the same throughout a spell
is especially problematic for higher-order spells where different
sex composition of previous births can lead to substantial differences 
both in the likelihood of progressing to the next birth and
how soon couples want their next child if they are going to have one.

The introduction of prenatal sex determination is likely to exacerbate 
any bias from the proportionality assumption.
First, sex-selective abortions affect birth spacing, and use
varies across groups.
Second, the use of sex selection may vary within a spell, which means 
that the effects of covariates vary within the spell as well.

In summary, my proposed model is a discrete time, non-proportional, competing risk 
hazard model with two exit states: either a boy or a girl is born.
For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is
\begin{equation}
 h_{ijt} = \Pr (T_i=t, J_i=j \mid T_i \geq t; \mathbf{Z}_{it},\mathbf{X}_{i} ),
\end{equation}
where $T_i$ is a discrete random variable that captures when woman $i$'s birth occurs.
To ease presentation the indicator for spell number is suppressed.
The explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, include 
individual, household, and community characteristics discussed below.

The hazard rate is specified as
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2
 \label{eq:hazard}
\end{equation}
where $D_{j}(t)$ is the piece-wise linear baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
where $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not place 
overly strong restrictions on the baseline hazard.

The explanatory variables in $\mathbf{Z}$ and the interactions between them 
are the non-proportional part of the model, which means that they are
interacted with the baseline hazard:
\begin{equation}
 \mathbf{Z}_{it} = D_j(t) \times (\mathbf{Z}_1 + Z_2 + \mathbf{Z}_1 \times Z_2),
\end{equation}
where $D_j(t)$ is the piece-wise linear baseline hazard and $\mathbf{Z}_1$ captures sex 
composition of previous children, if any, and $Z_2$ captures area of residence.
The non-proportionality allows the effects of the main explanatory variables on the probabilities 
of having a boy, a girl, or no birth to vary over time within a spell.
The remaining explanatory variables, $\mathbf{X}$, enter proportionally,
but to further minimize any potential bias from assuming proportionality, estimations 
are done separately for different levels of mothers' education and for different 
time periods.
% The exact specifications and the individual variables are described below.

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, transforming the data, so each observation is an interval---here equal
to three months---the model can be estimated using a standard multinomial logit model.%
\footnote{
A potential issue is that the multinomial model assumes that alternative 
exit states are stochastically independent,
also known as the Independence of Irrelevant Alternatives (IIA) assumption.
This assumption rules out any individual-specific unmeasured or 
unobservable factors that affect both the hazard of having a girl and the 
hazard of having a boy.
I, therefore, include a proxy for fecundity discussed in Section \ref{sec:data}.
Also, the multivariate probit model can be used as an alternative
to the multinomial logit because the IIA is not imposed \citep{han90}.
The results are substantially identical between these two models and
available upon request.
}
In the reorganized data the outcome variable is 0 if the
woman does not have a child in a given interval (the base outcome), 1 if 
she gives birth to a son in that interval, and 2 if she gives birth to 
a daughter in that interval.

The main downside of this set-up is that interpretation of the estimated 
coefficients is challenging;
the coefficients show the change in hazards relative to the base outcome, 
here no birth, rather than the hazard of an event.
Hence, a positive coefficient does not necessarily imply that the associated 
exit state becomes more likely as the explanatory variable increases because the 
probability of the other exit state may increase even more \citep{thomas96}.

It is, however, straightforward to calculate the predicted probabilities of 
having a boy and of having a girl for each $t$ within a spell, conditional on 
a set of explanatory variables and not having had a child before that period.
The predicted probability of having a boy in period $t$ for a given set of 
explanatory variable values, $\mathbf{Z}_k$ and $\mathbf{X}_k$, is
\begin{equation}
P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt}, t ) 
=  
\frac{ \exp(D_j(t) + \alpha_{1t}' \mathbf{Z}_{kt} + \beta_1' \mathbf{X}_{k} )}
{1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt} ' \mathbf{Z}_{kt} + \beta_l ' \mathbf{X}_{k})},
\label{eq:probability_boy}
\end{equation}
and the predicted probability of having a girl is
\begin{equation}
P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t ) 
=  
\frac{ \exp(D_j(t) + \alpha_{2t}'\mathbf{Z}_{kt} + \beta_2'\mathbf{X}_{k} )}
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{kt} + \beta_l'\mathbf{X}_{k})}.
\label{eq:probability_girl}
\end{equation}
Within each period the probability of not having a birth in period $t$
is $1-P(b_{t})-P(g_{t})$.

The distribution of spacing is captured by the survival curve, which shows 
the probability of not having had a birth yet by spell duration, for
a given set of explanatory variables.
The survival curve at time $t$ is 
\begin{equation}
S_{t} 
= 
\prod_{d=1}^t 
\left( 
	1- \left(P(b_{d} | \mathbf{X}_{k}, \mathbf{Z}_{kd}, d) 
	+ P(g_{d} | \mathbf{X}_{k}, \mathbf{Z}_{kd}, d) \right) 
\right),
 \label{eq:survival}
\end{equation}
or equivalently
\begin{equation}
S_{t} 
= 
\prod_{d=1}^t
\left(
\frac{ 1 }
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{ld}'\mathbf{Z}_{kd} + \beta_l'\mathbf{X}_{k})}
\right).
\end{equation}

An important issue is that the probability of ever having a next birth varies 
across groups.
For example, as the parity progression literature shows, couples with stronger 
son preference are less likely, for a given number of prior births, to ever have 
a next birth if they already have one or more sons.
Direct comparison of standard survival curves, therefore, tells us little 
about how the spread of sex selection affects birth spacing across groups.

To overcome this problem, I condition on the predicted likelihood of parity 
progression when examining birth spacing measures, such as the median duration 
to a birth.
How well this approach works depends on whether the spell length covered is 
sufficiently long that few women are likely to give birth after the spell cut-off.
I discuss the choice of spell length below.
It is important to note that the approach is not the same as merely calculating 
the birth spacing measures for women who already have a given 
parity child in the survey because that number does not take into account
the censoring of spells that will eventually lead to a birth.
In addition to calculating conditional birth spacing measures for different 
percentages, I also present graphs of both standard survival curves and
survival curves conditional on parity progression (which therefore
begin at 100\% and end at 0\%).

Finally, given the predicted probabilities of having a boy and of having 
a girl for each $t$ within a spell, is it easy to calculate the estimated 
percentage of children born that are boys, $\hat{Y}$, at each $t$  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100.
\label{eq:probability_son}
\end{equation}
Combining the percentage boys and the likelihood of exiting the spell 
across all $t$ gives the predicted percent boys born over the entire spell.
In addition to the predicted percent boys over the entire spell, 
I also present graphs of how the percentage boys born vary across time
within a spell, together with the associated confidence interval for given 
values of explanatory variables calculated using the Delta method.

>>>>

\date{November 2018}

@@

The unit of analysis is a spell, the period from one birth---or for the first spell from 
marriage---to the next birth.
The model is a discrete time, non-proportional, competing risk hazard model with two exit 
states: either a boy or a girl is born.
The standard approach in the spacing literature is to use a single-exit model, but
access to sex selection means that the sex of the next child born is not random 
and that the spacing to girls and boys depends on parents' choices.
Both the non-randomness and differential spacing can be captured with a competing risk set 
up.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

I use a non-proportional hazard specification, which allows the shape of the hazard 
functions to vary across groups, rather than the standard proportional hazard function 
for three reasons.
First, it is unlikely---even in the absence of prenatal sex determination---that the 
effect of household characteristics such as the sex composition of previous births have 
the same effect throughout the entire spell.%
\footnote{
Assuming that the effect of sex composition is the same throughout a spell is especially 
problematic for higher-order spells where different sex composition of previous births can 
lead to substantial differences both in the likelihood of progressing to the next birth and
how soon couples want their next child if they are going to have one.
}
Second, the introduction of prenatal sex determination likely exacerbates any bias 
from the proportionality assumption since sex-selective abortion use varies across groups
and affects birth spacing, and because a household's use of sex selection may vary 
within a spell, which means that the effects of covariates vary within the spell as well.
Finally, the use of a non-proportional specification also mitigates any potential effects 
of unobserved heterogeneity when used in conjunction with a flexible baseline hazard 
\citep{Dolton1995}.

For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is specified as
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2
 \label{eq:hazard}
\end{equation}
where the explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, include 
individual, household, and community characteristics discussed below,
and $D_{j}(t)$ is the piece-wise linear baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
with $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not place 
overly strong restrictions on the baseline hazard.

The explanatory variables in $\mathbf{Z}$ and the interactions between them 
are the non-proportional part of the model, which means that they are
interacted with the baseline hazard:
\begin{equation}
 \mathbf{Z}_{it} = D_j(t) \times (\mathbf{Z}_1 + Z_2 + \mathbf{Z}_1 \times Z_2),
\end{equation}
where $D_j(t)$ is the piece-wise linear baseline hazard and $\mathbf{Z}_1$ captures sex 
composition of previous children, if any, and $Z_2$ captures area of residence.
The remaining explanatory variables, $\mathbf{X}$, enter proportionally,
but to further minimize any potential bias from assuming proportionality, estimations 
are done separately for different levels of mothers' education and for different 
time periods.

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, transforming the data, so each observation is an interval---here equal
to three months---the model can be estimated using a standard multinomial logit model.%
\footnote{
The multinomial model assumes that alternative exit states are stochastically independent
(Independence of Irrelevant Alternatives or IIA), which rules out any individual-specific 
unmeasured or unobservable factors that affect both the hazard of having a girl and the 
hazard of having a boy.
I, therefore, include a proxy for fecundity discussed in Section \ref{sec:data}.
Also, the multivariate probit model can be used as an alternative because the IIA is not 
imposed \citep{han90}.
The results are substantially identical between these two models and
available upon request.
}
In the reorganized data the outcome variable is 0 if the woman does not have a child in a 
given interval (the base outcome), 1 if she gives birth to a son in that interval, and 2 
if she gives birth to a daughter in that interval.

The distribution of spacing is captured by the survival curve, which shows the probability 
of not having had a birth yet by spell duration, for a given set of explanatory variables.
The survival curve at time $t$ is 
\begin{equation}
\label{eq:survival}
S_{t} 
= 
\prod_{d=1}^t
\left(
\frac{ 1 }
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{ld}'\mathbf{Z}_{kd} + \beta_l'\mathbf{X}_{k})}
\right).
\end{equation}

Because the probability of ever having a next birth likely varies across groups, a direct 
comparison of standard survival curves tells us little about how the spread of sex selection 
affects birth spacing across groups.
I, therefore, condition on the predicted likelihood of parity progression when examining 
birth spacing measures, such as the median duration to a birth.
The reliability of this approach depends on whether the spell length covered is 
sufficiently long that few women are likely to give birth after the spell cut-off.
I discuss the choice of spell length below.%
\footnote{
It is important to note that the approach is not the same as merely calculating 
the birth spacing measures for women who already have a given 
parity child in the survey because that number does not take into account
the censoring of spells that will eventually lead to a birth.
}
In addition to calculating conditional birth spacing measures for different 
percentages, I also present graphs of both standard survival curves and
survival curves conditional on parity progression (which therefore
begin at 100\% and end at 0\%).

Interpretation of the model coefficients is challenging \citep{thomas96}.
It is, however, possible to calculate the predicted probabilities of 
having a boy and of having a girl for each $t$ within a spell, conditional on 
a set of explanatory variables and not having had a child before that period.
The predicted probability of having a boy in period $t$ for a given set of 
explanatory variable values, $\mathbf{Z}_k$ and $\mathbf{X}_k$, is
\begin{equation}
P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt}, t ) 
=  
\frac{ \exp(D_j(t) + \alpha_{1t}' \mathbf{Z}_{kt} + \beta_1' \mathbf{X}_{k} )}
{1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt} ' \mathbf{Z}_{kt} + \beta_l ' \mathbf{X}_{k})},
\label{eq:probability_boy}
\end{equation}
and the predicted probability of having a girl is
\begin{equation}
P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t ) 
=  
\frac{ \exp(D_j(t) + \alpha_{2t}'\mathbf{Z}_{kt} + \beta_2'\mathbf{X}_{k} )}
{1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{kt} + \beta_l'\mathbf{X}_{k})}.
\label{eq:probability_girl}
\end{equation}
Within each period the probability of not having a birth in period $t$
is $1-P(b_{t})-P(g_{t})$.

Given the predicted probabilities of having a boy and of having a girl for each 
$t$ within a spell, is it easy to calculate the estimated percentage of children born that 
are boys, $\hat{Y}$, at each $t$  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100.
\label{eq:probability_son}
\end{equation}
Combining the percentage boys and the likelihood of exiting the spell 
across all $t$ gives the predicted percent boys born over the entire spell.
In addition to the predicted percent boys over the entire spell, 
I also present graphs of how the percentage boys born vary across time
within a spell, together with the associated confidence interval for given 
values of explanatory variables calculated using the Delta method.
